---
title: "Transformation activity scores (TASs) for FGFR1-4 variants"
author: "Masachika Ikegami"
date: "5/15/2019"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
    number_sections: yes
    toc: true
  html_document:
    df_print: paged
    toc: yes
header-includes:
- \usepackage{amsmath,amssymb}
- \usepackage{mathspec}
- \setlength{\mathindent}{0pt} 
classoption: xelatex,ja=standard,fleqn 
---
\pagebreak

# Setting up
```{r setup, results='hide'}
Working = "home/TAS"

knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(error=FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=6, fig.align='center', fig.pos='H', out.extra='', tidy.opts = list(width.cutoff = 45))
knitr::opts_knit$set(root.dir=Working)
setwd(Working)

options(scipen=100)
library(stringr)
library(rstan)
library(tidyverse)
library(reshape2)
library(shinystan)
library(codetools)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

# Loading the experimental results of focus formation assay and low-serum cell proliferation assay
```{r}
filename = "Data.csv"
v_list = NULL
N = NULL
V = NULL
I = 4
FFA_Data = read.csv(file = filename, stringsAsFactors=F)
FFA_Data$Name = str_replace_all(FFA_Data$Name, pattern="\\*", replacement="STOP")
FFA_Data$AA = str_sub(FFA_Data$Name, 1,-3)
v_l = unique(FFA_Data$AA)
anchor = "FGFR1_WT"
v_l2 = v_l[-which(v_l %in% anchor)]
v_list =　c(anchor,v_l2)
N = length(FFA_Data$score)
V = length(v_list)
var_num = rep(0,N)
for(k in 1:N){
  for(j in 1:V){
    if(FFA_Data$AA[k] == v_list[j]){
      var_num[k] = j
    }
  }
}
FFA_Data = cbind(FFA_Data, var_num)
```

# Ordered logit model simulation code in Stan language
```{r}
stan_code_F <- "
data{
  int I;            // Number of classes
  int N;            // Number of measurements
  int B[N];         // Batches of each measurement
  int BN;           // Number of the batches
  int V;            // Number of the variants
  int Y[N];         // Results of each measurement 
  int variant[N];   // Variant ID of each 
}

parameters{
    ordered[I-1] d;   // Thresholds for the ordered logit model
    simplex[V] eta;      // Functional scores of each variant
    vector[BN] be_t;    // Batch-specific random effects, intercept
    vector<lower=0>[BN] tau;  // Batch-specific random effects, slope
}
transformed parameters{
    vector[BN] bet;
    bet[1] = -tau[1]/V;
    for(b in 2:BN){
      bet[b] = be_t[b];
    }
}
model{
  for(n in 1:N){
    Y[n] ~ ordered_logistic(bet[B[n]]+tau[B[n]]*eta[variant[n]], d);
  }
}

generated quantities{
  vector[V] Delta_Eta;
  for(v in 1:V){
    Delta_Eta[v] = eta[v]-eta[1];
  }
}

"
write(stan_code_F, file="stan_code_FGFR.stan")
stanmodel_F <- stan_model(file="stan_code_FGFR.stan")
```

## Execution of Hamiltonian Monte Carlo simulation
Change "chains", "iter", "warmup" referring to R-hat values and posterior expectations. 
```{r,eval=FALSE}
fit_FFA = NULL
HMC_data <- list(N = N, V = V, I = I,
                  Y = FFA_Data$score, variant = FFA_Data$var_num, 
                  B = FFA_Data$batch, BN = max(FFA_Data$batch))
fit <- stan(file="stan_code_FGFR.stan", data=HMC_data,
            control=list(adapt_delta=0.9999, max_treedepth = 20),
            seed=1234, chains=4, iter=3500, warmup=1500, thin=2)
fit_FFA = fit
```

# Quality check 
Judge the convergence of simulation if R-hat values of all parameters are < 1.1 or not.
```{r eval=FALSE}
launch_shinystan(fit_FFA)
```

## Simulation summary 
```{r}
DF = NULL
DF_eta = NULL
DF_d = NULL
df_eta <- subset(summary(fit_FFA,probs = c(0.025, 0.05, 0.50, 0.95, 0.975))$summary, grepl("eta",
                rownames(summary(fit_FFA,probs = c(0.025, 0.05, 0.50, 0.95, 0.975))$summary)))
df_Eta <- subset(summary(fit_FFA,probs = c(0.025, 0.05, 0.50, 0.95, 0.975))$summary, grepl("Eta",
                rownames(summary(fit_FFA,probs = c(0.025, 0.05, 0.50, 0.95, 0.975))$summary)))
df_d <- subset(summary(fit_FFA,probs = c(0.025, 0.05, 0.50, 0.95, 0.975))$summary, grepl("d",
                rownames(summary(fit_FFA,probs = c(0.025, 0.05, 0.50, 0.95, 0.975))$summary)))
df_beta <- subset(summary(fit_FFA,probs = c(0.025, 0.05, 0.50, 0.95, 0.975))$summary, grepl("bet",
                rownames(summary(fit_FFA,probs = c(0.025, 0.05, 0.50, 0.95, 0.975))$summary)))
df_tau <- subset(summary(fit_FFA,probs = c(0.025, 0.05, 0.50, 0.95, 0.975))$summary, grepl("tau",
                rownames(summary(fit_FFA,probs = c(0.025, 0.05, 0.50, 0.95, 0.975))$summary)))
df_ <- rbind(df_eta,df_Eta,df_d,df_tau,df_beta)
DF = df_
DF_eta = df_eta
DF_d = df_d
```

## Sampling from posterior distribution and estimating TAS
Draw 500 parameter samples from the trace randomly, then generate  values, taking batch-to-batch ratio into account. 
```{r}
FFA_result = NULL
len <- length(v_list)
eta <- NULL
eta$ave <- DF_eta[1:len]
eta$name <- v_list
repN <- 500
BN <- max(FFA_Data$batch)
BN_ <- NULL
for(j in 1:BN){
  BN_ <- c(BN_,length(subset(FFA_Data, batch == j)$batch))
}
NY <- length(v_list)
p_Y <- array(0,c(repN, BN, NY))
d.ext<-rstan::extract(fit_FFA)
for(k in 1:repN){
  for(j in 1: NY){
    tmp <- 0
    for(l in 1: BN){
      p_Y[k,l,j] = p_Y[k,l,j]  + (d.ext$bet[k,l] + d.ext$tau[k,l]*d.ext$eta[k,j]) * BN_[l] * BN / sum(BN_)
    }
  }
}
eta$le95 <- DF_eta[(len*3+1):(len*4)]
eta$ue95 <- DF_eta[(len*7+1):(len*8)]
y_ave <- rep(0,len)
y_class <- rep(4,len)

for(j in 1:len){
  y_ave[j] <- mean(p_Y[,,j])
  eta$ue95[j] <- t.test(p_Y[,,j])[4]$conf.int[2]
  eta$le95[j] <- t.test(p_Y[,,j])[4]$conf.int[1]
  if(y_ave[j] < DF_d[3])
      y_class[j] <- 3
  if(y_ave[j] < DF_d[2])
      y_class[j] <- 2
  if(y_ave[j] < DF_d[1])
      y_class[j] <- 1
}
eta$y_ave <- y_ave
eta$y_class <- y_class
filename <- "FGFR-FFA_score.tsv"
write.table(eta, file=filename, sep='\t',
          quote=FALSE, col.names=NA)
```
